---
title: "Project_1"
author: "Owen Gould"
date: "2024-02-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(plotly)
df <- read.csv("https://grinnell-statistics.github.io/sta-310-s24/projects/USStates.csv")
```

```{r}
# Add `CovidMortalityRate` variable to our data frame
df <- df %>%
  mutate(CovidMortalityRate = (deaths / cases) * 100)
mod <- lm(CovidMortalityRate ~ Insured + HighSchool + HouseholdIncome, data = df)
summary(mod)
plot(mod)
```
Fitted mode: $\hat{y}=8.746654 + 0.056223\text{Insured} -0.136215\text{HighSchool} + 0.005709\text{HouseoldIncome}$.
Multiple R-squared = $0.3034$ (low). *Variability not super well-explained by linear relationship between resposne and predictors*.
`Insured` significant (p-value of around $0.001$), `HighSchool` highly significant (p-value of nearly $0$), `HouseholdIncome` not significant (p-value of around $0.39$) *Shows which covariates are statistically significant*.
F-statistic: 6.667 on 3 and 46 DF with associated p-value of nearly zero. *Overall model is strong*.
Model assumptions feel fair (move plots to Appendix).
Check assumptions (of equal variance, etc.) individually?
Surprising that `Insured` and `HouseholdIncome` were instead positively associated with covid mortality rate and that `HighSchool` was *negatively* associated with covid mortality rate.
Generalizations: if socioeconomic factors predict health outcomes/response to Covid, generalize to other pandemics?
Limitations/concerns: `HoldholdIncome` measured with mean instead of median (mean more sensitive to outliers); one data point in time (pandemic and response to it changed over time); reporting of data (Covid morality rate calculations)?
*Find out if zero is in 95% confint for slope coefficients*
Further Research: why were the results what we expected?
